### 2013-06-07 inst/examples/BUGS/allen.Rmd
different mcmc conditions

```{r stateeq}
f <- RickerAllee
p <- c(1, 10, 5)
K <- 10  # approx, a li'l' less
allee <- 5 # approx, a li'l' less
```

@@

```{r mle, dependson="obs"}
set.seed(12345)
estf <- function(p){ 
    mu <- f(obs$x,0,p)
    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
}
par <- c(p[1]+abs(rnorm(1,0,1)), 
         p[2]+rnorm(1,0,.1), 
         p[3]+rnorm(1,0, .1), 
         sigma_g + abs(rnorm(1,0,.1)))
o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
f_alt <- f
p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
sigma_g_alt <- as.numeric(o$par[4])

@@

```{r gp_traces_densities, dependson="gp"}
plots <- summary_gp_mcmc(gp)
```

@@

```{r jags-setup}
y <- obs$x[-1] 
N <- length(y);
jags.data <- list("N","y")
n.chains <- 2
n.iter <- 1e5
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
n.update <- 10
```

@@

```{r common-priors}
stdQ_prior_p <- c(0.0001, 100)
stdR_prior_p <- c(1e-4, .1)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
```

@@

jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0,.1),
       "logr0"=log(rlnorm(1,0,.5) ),
       "logtheta"=log(2 * rlnorm(1,0,.1)), 
       "stdQ"= 0.2 * rlnorm(1,0,.1),
       "stdR"= 1e-5 * rlnorm(1,0,.1),
       "x"=y,        
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

@@

```{r ricker-mcmc, dependson="ricker-model", results="hide"}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0,.5),
       "logr0" = log( rlnorm(1,0,.5) ),
       "stdQ"= 0.2 * rlnorm(1,0,.2),
       "stdR"= 1e-5 * rlnorm(1,0,.1),
       "x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- do.call(jags.parallel, 
                       list(data=jags.data, inits=jags.inits, 
                            jags.params, n.chains=n.chains, 
                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                            model.file="ricker.bugs"))
```

@@

```{r myers-output, dependson="myers-traces"}
myers_pardist <- acast(myers_posteriors[2:3], 
                        1:table(myers_posteriors$variable) ~ variable, 
                        subset=.(variable!="deviance")) 
myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c("r0", "theta", "K")])
head(myers_pardist)
myers_bayes_pars

@@

```{r deviances, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
allen_deviance <- pardist[,'deviance']
ricker_deviance <- ricker_pardist[,'deviance']
myers_deviance <- myers_pardist[,'deviance']
true_deviance <- 2*estf(c(p, sigma_g))
mle_deviance <- 2*estf(c(est$p, est$sigma_g))

>>>>

```{r stateeq}
f <- RickerAllee
p <- c(2, 8, 5)
K <- 10  # approx, a li'l' less
allee <- 5 # approx, a li'l' less
```

@@

```{r mle, dependson="obs"}
set.seed(12345)
estf <- function(p){ 
    mu <- f(obs$x,0,p)
    -sum(dlnorm(obs$y, log(mu), p[4]), log=TRUE)
}
par <- c(p[1]*rlnorm(1,0,.4), 
         p[2]*rlnorm(1,0,.3), 
         p[3]*rlnorm(1,0, .3), 
         sigma_g * rlnorm(1,0,.3))
o <- optim(par, estf, method="L", lower=c(1e-5,1e-5,1e-5,1e-5))
f_alt <- f
p_alt <- c(as.numeric(o$par[1]), as.numeric(o$par[2]), as.numeric(o$par[3]))
sigma_g_alt <- as.numeric(o$par[4])

@@

```{r gp_traces_densities, dependson="gp"}
plots <- summary_gp_mcmc(gp, burnin=1e4, thin=300)
```

@@

```{r jags-setup, dependson="obs"}
y <- obs$x[-1] 
N <- length(y);
jags.data <- list("N","y")
n.chains <- 2
n.iter <- 40000
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
n.update <- 10
```

@@

```{r common-priors}
stdQ_prior_p <- c(1e-6, 100)
stdR_prior_p <- c(1e-6, .1)
stdQ_prior  <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior  <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])
```

@@

jags.inits <- function(){
  list("K"= 8 * rlnorm(1,0,.1),
       "logr0"=log(2 * rlnorm(1,0,.1) ),
       "logtheta"=log(5 * rlnorm(1,0,.1)), 
       "stdQ"= 0.05 * rlnorm(1,0,.1),
       "stdR"= 1e-5 * rlnorm(1,0,.1),
       "x"=y,        
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

@@

```{r ricker-mcmc, dependson="ricker-model", results="hide"}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"= 10 * rlnorm(1,0,.2),
       "logr0" = log( rlnorm(1,0,.2) ),
       "stdQ"= 0.01 * rlnorm(1,0,.1),
       "stdR"= 1e-5 * rlnorm(1,0,.1),
       "x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- do.call(jags.parallel, 
                       list(data=jags.data, inits=jags.inits, 
                            jags.params, n.chains=n.chains, 
                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                            model.file="ricker.bugs"))
```

@@

```{r myers-output, dependson="myers-traces"}
A <- myers_posteriors
A$index <- A$index + A$chain * max(A$index) # Combine samples across chains by renumbering index 
myers_pardist <- acast(A, index ~ variable)
### myers_pardist <- acast(myers_posteriors[2:3], 1:table(myers_posteriors$variable) ~ variable) 
myers_pardist[,"logK"] = exp(myers_pardist[,"logK"]) # transform model parameters back first
myers_pardist[,"logr0"] = exp(myers_pardist[,"logr0"]) # transform model parameters back first
myers_pardist[,"logtheta"] = exp(myers_pardist[,"logtheta"]) # transform model parameters back first
colnames(myers_pardist)[colnames(myers_pardist)=="logK"] = "K"
colnames(myers_pardist)[colnames(myers_pardist)=="logr0"] = "r0"
colnames(myers_pardist)[colnames(myers_pardist)=="logtheta"] = "theta"

bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates
myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c("r0", "theta", "K")])
head(myers_pardist)
myers_bayes_pars

@@

```{r deviances, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
allen_deviance <- posterior.mode(pardist[,'deviance'])
ricker_deviance <- posterior.mode(ricker_pardist[,'deviance'])
myers_deviance <- posterior.mode(myers_pardist[,'deviance'])
true_deviance <- 2*estf(c(p, sigma_g))
mle_deviance <- 2*estf(c(est$p, est$sigma_g))
